---
title: "Snowshoe Hares Exploratory Report"
author: "Alison Sells"
date: "11/18/2020"
output: html_document
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(dplyr)
library(janitor)
library(ggbeeswarm)
library(lubridate)
library(ggpubr)
library(broom) 
library(kableExtra)
library(effsize)


```

### 1.Introduction

This report provides and exploratory overview of differences in juvenile hare weights between from three grid sites in the Bonanza Creek Experimental Forest from 1999 to 2012(Kielland et. al 2017). Specifically, weights, and hind lengths are explored and compared across the male, female and unidentified hares. To explore this dataset multiple methods of visualization are used in preparation to run statistical tests to determine the correlation between the various variables that are being observed. The study takes place in the Bonanza Creek Experimental Forest which is in Fairbanks Alaska.   

### 2.Data and Analyses
This data was collected from the Bonanza Creek Experimental Forest in Fairbanks Alaska form 1999 to 2012. It was collected by using three grid sites to see the variation of many variables including weight, hindfoot length, as well as the concentration of hares captured at different grid sites. This report will use various methods of visualization including a histograms,boxplots, and tables; and will utilize t tests, and a linear regression model to look at significance levels and residual impacts. All analyses are in R version 4.0.2 using RStudio version 1.3.1056.

### 3.Exploratory Findings
#### A. Annual juvenile hare trap counts
Juvenile hare trappings were totaled annually from 1999 to 2012
```{r, echo = FALSE, message = FALSE, warning = FALSE}
hares <- read.csv(here("data","bonanza_hares.csv"))
```
```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Data wrangling to get only juvenile hares
j_hares <- hares %>% 
 filter(age== "j")

 # Data wrangling to separate year, and create a count
hares_count<-j_hares %>% 
  mutate(date_new= mdy(date)) %>% 
  mutate(year_new= year(date_new)) %>% 
  count(year_new)

## Plotting the number of juvenile hare trapped per year
ggplot()+ geom_col(data = hares_count, aes(x= year_new, y= n), fill= "cadetblue4")+
  scale_x_continuous(breaks = c(1999, 2001, 2003, 2005, 2007, 2009, 2011))+
  labs(x= "Year", y= "Number of juvenile hare trappings")+
  theme_minimal()
```


**Figure 1.** Graph detailing the total number of hare trappings within the Bonanza Creek Experimental Forest during 1999-2012,Fairbanks Alaska from Kielland et. al(2017)

#### Major Takeaways
  When looking at the total number of snowshoe hare trappings at the site, one can see that the maximum trappings was in 1999 with 126 trappings, and the lowest was in 2010 with 2 trappings. The mean number of trappings was 31.5 per year with a median value of 18.5 trappings per year. This data also shows an overall trend where the number of juveniles trapped decreased significantly throughout the years. However, looking forward to something to consider is not just the year captured but the season and day as well as correlate are the trappings to each trap itself to see where the distribution of trappings are happening. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#summmarize juvenile trappings
juvenile_trappings_sum <-hares_count %>%
  group_by(year_new) %>%
  summarize(
    mean_trappings = round(mean(n, na.rm = TRUE),2),
    median_trapping = round(median(n, na.rm = TRUE),2),
    )
```



#### B. Visualize juvenile hare weights
The following visualization aims to explore the relationships between juvenile hare weights as related to sex and the grid site of capture. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Pull vectors of weight for male and female hares by site

m_hares <- j_hares %>% 
  filter(sex== "m") %>% 
 pull(weight)  

f_hares <- j_hares %>% 
  filter(sex== "f") %>% 
 pull(weight) 

#Keeping only the necessary columns before making a Beeswarm
hares_sex_site <-j_hares %>% 
  select(sex, weight, grid)

# Beeswarm
ggplot(data = hares_sex_site, aes(x= sex, y= weight))+
  geom_beeswarm(aes(color= sex))+
  facet_wrap(~grid)+
  geom_boxplot(fill= NA, width = 0.3, outlier.color = NA)+
  stat_summary(fun = mean,
               geom = "point",
               shape= 20,
               size= 3,
               color= "black",
               fill= "black")+
  labs(title = "Comparison of Male and Female weights in Juvenile Snowshoe Hares", x= "Sex", y="Weight(grams)")
```


**Figure 2.** Hare weight observations (grams) for  male, female and unidentified juvenile snowshoe hares from sites Bonanza Balck Spruce (bonbs),Bonanza Mature(bonmat), and Bonanza Riparian(bonrip).Coral (female), Teal(male), and Gray(N/A) represent individual observations for a juvenile hare's weight. Box endpoints indicate the 25th and 75th percentile values; the black line and black point within the box indicate the median and mean value for each sex, respectively.Data: Kielland et al(2017)

#### Major Takeaways
  When looking at the comparison of juvenile weights between male and female hares at the various sites, one can see that the males tend to be more centrally distrusted in that there is less overall spread between the heaviest juvenile male and the lightest juvenile male when compared to the spread of the juvenile female hares and those whose sex was not identified. It is interesting to see how the hares from Bonanza Riparian   were on average much heavier than then hares from Bonanza Mature. 

#### C. Juvenile weight comparison between male and female snowshoe hares
The following will be a deeper exploration of the Juvenule weight comparison within the snowshoe hare sample
```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Summarize
juvenile_weight_sum <- hares_sex_site %>%
  group_by(sex,grid) %>%
  summarize(
    mean_weight = round(mean(weight, na.rm = TRUE),2),
    sd_weight = round(sd(weight,na.rm = TRUE),2),
    sample_size = n())

```


**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for three grid sites of Juvenile Snowshoe Hares. Data: Kielland et. al(2017)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
juvenile_summary_table<- juvenile_weight_sum %>% 
  select(sex, mean_weight,sd_weight,sample_size) %>% 
  group_by(sex)

juvenile_summary_table %>% 
  kable(col.names = c("Sex", "Mean weight (grams)", "Standard deviation (grams)", "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

```{r, include=FALSE}

#Graph 

# Plot histograms & qq-plots to check assumptions
hist(m_hares) #Looks mostly normal
qqnorm(m_hares)# Pretty Linear
hist(f_hares) # Possibly a bit negatively skewed
qqnorm(f_hares)#Pretty Linear

# Overall: assumptions of normality hold, & sample size additionally >> Central Limit Theorem (means will be normally distributed). Two-sample t-test OK for means comparison! 


#t-test
juv_weight_ttest <- t.test(m_hares, f_hares)
juv_weight_ttest

#Cohen's d test to find effect size
cohen.d(m_hares, f_hares, na.rm = TRUE)
```
#### Major Takeaways

After running a t test and a chon’s d test, the difference in mean weights between male and female hares is approximately 90 grams with a cohen’s D estimate of 0.29. The outcome of the two sample t tests revealed a p value of 0.007 which indicates a very low probability that these samples were taken with populations with the same mean. Although p values are not the stand alone metric of significance, that combined with the low Cohen’s D score  help lead to a statement that it is extremely unlikely that these samples came from populations with the same mean weights. 


#### D. Relationship between juvenile weight and hind foot length
The following will explore the relationship between juvenile weight and hind foot length in the snowshoe hare population
```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Exploratory visualization
juveniles_hind <- j_hares %>%
  select(hindft,weight)
 

  ggplot(data= juveniles_hind, aes(x= hindft, y= weight))+
  geom_point() +
    labs(title ="Comparison of juvenile Snowshoe Hare weight and hindfoot length",x= "hind foot length(mm)", y= "weight(grams)")
```


**Figure 3.**  Comparison of juvenile hare weights and hinndfoot measurements from all of the juvenile hares within the sample.Each dot represents an individual observation for a juvenile hare's weight and hindfoot measurement. Data: Kielland et al(2017)
```{r include=FALSE}
#Linear model
hare_lm <- lm(hindft~weight, data =juveniles_hind)

summary(hare_lm)

```


#### Linear Regression
The following will be a linear regression of the juvenile hares data to test further the relationship between hindfoot length and weight.
```{r include=FALSE}
#Plot linear model
plot(hare_lm)
```
```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data = juveniles_hind, aes(x = hindft, y = weight)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  labs(x= "hind foot length(mm)", y= "weight(grams)")+
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700)
```




**Figure 4.**  Linear Regression model Comparing juvenile hare weights and hinndfoot measurements from all of the juvenile hares within the sample.Each dot represents an individual observation for a juvenile hare's weight and hindfoot measurement. Data: Kielland et al(2017)

With a R squared value of .2959, this indicates that 29 percent on the juvenille's hare weight is determined by their hindfoot length in millimeters. However, this figure is a little concerning when one considers the assumptions needed for the linear regression to show the data. The assumption of the data being relatively linear is concerning because of the concentration of data points between 100 and 150 mm and the lack of data where the hares' hindfoot length is less than 100 mm. ALso this model simplifies the relationship between tese two variables, and does not consider anything else that could describe the relationship.


### 4. Summary
Overall looking at the various visualizations revealed the sites where more hares were captured along with the relationship between the various factors including the hare's sex . These findings include:
- the most hares were captured during the beginning of the study in 1999 and then there was a decline of trappings from 2000 to 2012
- The hares from the Bonanza Riparian grid site were on average the heaviest, with the hares from the grid site being the lightest
- There is a significance difference in the means between the male and female juvenille hares indicating that is very unlikely that these samples were taken from populations where the means are the same
- looking at the linear regression brings some conerns especially since some of the assumptions were not met.

### 5. Citations
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14












