---
title: "Snowshoe Hares Exploratory Report"
author: "Alison Sells"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(dplyr)
library(janitor)
library(beeswarm)
library(lubridate)
library(ggpubr)
library(broom) 
library(kableExtra)
library(effsize)


```

#1.Introduction

This report 












#2.Data and Analyses



#3.Exploratory Findings
##A.Annual juvenile hare trap counts

```{r}
j_hares <- read.csv(here("data","bonanza_hares.csv"))
```



##B.Visualize juvenile hare weights


##C. Juvenile weight comparison between male and female snowshoe hares)




##D. Relationship between juvenile weight and hind foot length




#4. Summary



#5. Citations
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14












